Evolution of pre-copulatory and post-copulatory inbreeding behaviour

### 2016-09-27 ms_pre_post.Rmd
Edit Discussion, tweaks to results

---
title: 'Evolution of pre-copulatory and post-copulatory inbreeding behaviour'
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

<!--- Need to do 40k generations to see allele values level off, 40 replicates each:

Figure    1: 100 hours total
Figures   2: 48  hours total
Figures   3: 48  hours total
Figures   4: 48  hours total
Figures   5: 48  hours total

Nah, these are too much. I've decided to increase to 40k generations for the costs only (100 hours total). Reasoning for this is that when we look at the fixed traits, we are really interested in whether or not the initial invasion changes, hence fewer generations are better. When we start everything at zero and let it evolve, then long-term stuff becomes more important, so 40k are useful.

--->

@@

We first test our hypothesis that a focal female's ability to express post-copulatory choice with respect to kinship causes selection for polyandry by comparing simulations in which post-copulatory inbreeding strategy is fixed to random mating with respect to kinship to simulations in which tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy can all evolve. The left and right columns of Figure 1 show results for fixed versus evolving post-copulatory choice, respectively, across four rows of costs of tendency for polyandry. For low costs of tendency for polyandry, $c_{P} < 0.005$, selection caused expected tendency for polyandry allele values ($P_{a}$) to be positive only when post-copulatory inbreeding avoidance evolved (Figure 1A-D), and for all $c_{P}$, $P_{a}$ values were higher when post-copulatory inbreeding avoidance evolved than when it could not evolve. When post-copulatory inbreeding avoidance could not evolve, $P_{a}$ values were negative even in the absence of a direct cost ($c_{P}=0$; Figure 1A) because females that expressed polyandry were more likely to increase their inbreeding due to sampling males without replacement [@Duthie]. Strong post-copulatory inbreeding avoidance consistently evolved in all simulations, even when $P_{a}$ values were not expected to be positive (Figure 1F,H), because of the threshold nature of the polyandry phenotype; Figure 2 shows how phenotypic polyandry was common even when mean polyandry allele values within a populatoin were expected to be negative. In all simulations, pre-copulatory inbreeding avoidance also evolved ($M_{a}<0$), meaning that both pre- and post-copulatory inbreeding avoidance was predicted, and evolution of adaptive polyandry occurred given sufficiently low costs of tendency for polyandry.

@@

Figure 1 shows changes in mean allele values of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy across all eight possible combinations of costs. In Figure 1A, when no phenotypes are costly, inbreeding mating strategy and inbreeding fertilisation strategy evolve to inbreeding avoidance ($M_{a}<0$ and $F_{a}<0$), and females evolve to be polyandrous ($P_{a}>0$). Costs of phenotypes had a direct effect on evolution, causing inbreeding mating strategy and inbreeding fertilisation strategy phenotypes to evolve to values closer to zero (because realisation of inbreeding strategy costs increases with phenotype absolute values) and tendency for polyandry phenotypes to evolve to values that were more negative (because realisation of polyandry costs increases with positive phenotype values).

We can use Figure 1 to test the hypothesis that inbreeding mating strategy will evolve to inbreeding avoidance ($M_{a}<0$), but inbreeding fertilisation strategy will be negligible ($F_{a} \approx 0$) when mating costs are low ($c_{M}=0$) relative to the costs of tendency for polyandry and inbreeding fertilisation strategy ($c_{P}=0.02$ or $c_{F}=0.02$) by comparing different cost combinations to the null model in which all costs equal zero (Figure 1A). Figure 1A,C,D,G show that inbreeding avoidance in mating ($M_{a}<0$) evolved consistently whenever $c_{M}=0$. When inbreeding fertilisation strategy was costly and inbreeding avoidance strategy was not costly (Figure 1D,G), inbreeding fertilisation strategy initially evolved toward inbreeding avoidance ($F_{a} < 0$), but long-term evolution of inbreeding fertilisation strategy stabilised at random mating ($F_{a} \approx 0$). However, when only tendency for polyandry was costly (Figure 1C), both inbreeding mating strategy and inbreeding fertilisation strategy evolved to inbreeding avoidance over generations ($M_{a} < 0$ and $F_{a} < 0$).

We can also use Figure 1 to test the hypothesis that inbreeding fertilisation strategy will evolve to inbreeding avoidance ($F_{a}<0$), but inbreeding mating strategy will be negligible ($M_{a} \approx 0$) when tendency for polyandry and inbreeding fertilisation strategy costs are low ($c_{P}=0$ and $c_{F}=0$) relative to the cost of inbreeding mating strategy ($c_{M}=0.02$). Figure 1B shows that inbreeding avoidance through fertilisation bias evolved ($F_{a}<0$) under these conditions. While inbreeding mating strategy initially evolved toward inbreeding avoidance ($M_{a} < 0$), long-term evolution of inbreeding mating strategy stabilised at random mating ($M_{a} \approx 0$).

@@

Figure 4 shows change in inbreeding mating strategy allele values over generations when tendency for polyandry and inbreeding fertilisation strategy alleles were fixed for adaptive values of inbreeding avoidance $P_{a}=1$ and $F_{a}=-5$. Despite that females in these simulations could already avoid inbreeding after mating occurred, inbreeding mating strategy evolved towards inbreeding avoidance given costs of $c_{M}=0$ (Figure 4A) and $c_{M}=0.02$ (Figure 4B). We therefore did not find evidence that the existence of adaptive inbreeding avoidance through fertilisation bias in any way precluded the evolution of adaptive inbreeding avoidance through mate choice.

Figure 5 shows change in tendency for polyandry and inbreeding fertilisation strategy allele values over generations when inbreeding mating strategy alleles were fixed for adaptive inbreeding avoidance $M_{a}=-5$. Existence of adaptive inbreeding avoidance through mate choice precluded evolution of polyandry for all combinations of costs (Figure 5), meaning that evolution of adpative polyandry to avoid inbreeding through inbreeding fertilisation strategy was never predicted in populations where females were already avoiding inbreeding through mate choice. Further, while inbreeding fertilisation strategy evolved toward inbreeding avoidance ($F_{a}<0$) in the absence of costs (Figure 5A,B), allele values of inbreeding fertilisation strategy were not as negative as when adaptive inbreeding avoidance through mate choice was not fixed at the outset (Figures 1A,B,C,D). We therefore found strong evidence that the existence of adaptive inbreeding avoidance through pre-copulatory mate choice precluded evolution of adaptive inbreeding avoidance through post-copulatory inbreeding avoidance through fertilisation bias.

@@

Inbred offspring are typically less viable than outbred offspring, and when such inbreeding depression is strong, it is predicted to cause selection for inbreeding avoidance [Duthie and Reid *in press*; @Parker1979; @Parker2006; @Pusey1996; @Kokko2006]. In animals, inbreeding avoidance can be realised through both pre-copulatory and post-copulatory mechanisms [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, the conditions under which inbreeding avoidance is predicted to evolve via pre-copulatory, post-copulatory, or both pre- and post-copulatory mechanisms has not been modelled. There is therefore no framework for predicting what mechanisms of inbreeding avoidance are most likely to evolve in empirical populations where risk of fitness loss caused by inbreeding is high. Importantly, evolution of post-copulatory mechanisms of inbreeding avoidance requires females to mate multiply, meaning that post-copulatory inbreeding avoidance cannot occur unless polyandry has evolved in a population. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency to engage in polyandry, inbreeding mating strategy (pre-copulatory), and inbreeding fertilisation strategy (post-copulatory). Using *in silico* experiments, we tested four specific hypothesis to predict evolution of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy given diverse costs and evolutionary constraints. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of inbreeding fertilisation strategy, (2) selection was strong for low cost, and weak but not negligible for high cost, mechanisms of inbreeding avoidance, (3) constraints on evolution of pre-copulatory inbreeding avoidance greatly strengthened selection for post-copulatory inbreeding avoidance, but not vice versa, and (4) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of post-copulatory inbreeding avoidance, but not vice versa. More generally, we provide a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms.

Evolution of polyandry to avoid inbreeding
----------------------------------------------

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry to avoid inbreeding is predicted to be highly restrictive, requiring strong constraints on the availability of initial compared to additional mates, or the conditional expression of polyandry such that females only engaged in polyandry when their initial choice of mate was of low quality. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express choice after mating had occurred. In contrast, we found that when females could express choice after mating through biased fertilisation, selection for polyandry was greatly strengthened (Figure 1A), especially when constraints restricted evolution of inbreeding avoidance through mate choice (Figure 3). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g. @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. For example, @Pizzari2004 found that the number of sperm arriving to eggs is lower for full siblings of female red jungefowl (*Gallus gallus*) than it is for non-relatives, even after controlling for effects of order of oviposition and social familiarity. Similarly, Firman and Simmons [-@Firman2008; -@Firman2015] have shown evidence of post-copulatory inbreeding avoidance in house mice, finding fertilisation to be biased toward unrelated males independent of the order in which mating occurs [@Firman2008], perhaps due to post-copulatory mechanisms expressed through the secretion of gametic proteins originating in female's eggs [@Firman2015]. Interestingly, female house mice have also been shown to express pre-copulatory inbreeding avoidance  [@Potts1991; @Roberts2003]. Our results suggest that post-copulatory inbreeding avoidance is unlikely to evolve when females can already avoid inbreeding through pre-copulatory mechanisms (Figure 5), meaning that evolution of post-copulatory inbreeding avoidance could reasonably be hypothesised to have preceded evolution of pre-copulatory inbreeding avoidance in mice.

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found that female *T. commodus* are capable of controlling the transfer of sperm to their spermatheca, most likely through physical control of the spermathecal duct, and can reduce sperm storage despite being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1) or constrained (Figure 3).

Costs of polyandry and inbreeding avoidance
----------------------------------------------

Our results show that direct negative selection on phenotypes that increase female promiscuity or choosiness with respect to kinship decrease the strength of selection for tendency for polyandry and inbreeding avoidance (Figure 1). Importantly, the cost of tendency for polyandry only decreased female fitness given tendency for polyandry phenotype values greater than zero ($P_{p}>0$), but the threshold nature of the tendency for polyandry phenotype only resulted in selection against positive tendency for polyandry allele values ($P_{a}$) when the sum over a focal female's $20$ $P_{a}$ alleles was greater than zero because summations less than zero all resulted in equivalent monandrous phenotypes ($P_{p}=0$). Hence, even when tendency for polyandry was costly, selection for more negative $P_{a}$ values weakened as the frequency of negative $P_{a}$ values increased because each individual $P_{a}$ then became less influential in causing females to be monandrous [@Duthie]. This general property of threshold traits causes maladaptive phenotypes to frequently reappear within populations due to the mutation and recombination of continuous genetic values underlying discrete phenotypes [@Roff1996; @Roff1998]. Assuming that there is continuous genetic variation underlying threshold polyandry, polyandrous phenotypes are therefore sometimes predicted to persist within populations even when costly [@Duthie], potentially creating the possibility for selection on inbreeding fertilisation strategy when polyandry is expressed. This is most evident in Figure 1C, where mean allele values for polyandry are negative (red line), but strong post-copulatory inbreeding avoidance is still predicted (grey line). Consequently, costly polyandry does not necessarily preclude evolution of adaptive post-copulatory inbreeding avoidance in populations.

A low relative cost for one inbreeding strategy (e.g., pre-copulatory inbreeding avoidance) precluded a more costly complementary inbreeding strategy from persisting in a focal population (Figure 1), meaning that long-term persistence of both pre- and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of tendency for polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of inbreeding mating strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of inbreeding fertilisation strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations. <!--- add Tuni citation for depletion of energy? --->

@@

Constraints on polyandry and inbreeding avoidance
----------------------------------------------

@@

Constraints on pre-copulatory and post-copulatory inbreeding avoidance might vary widely in empirical systems. Female mate choice might be evolutionarily constrained if females are unable to discriminate among kin and therefore incapable of choosing to avoid inbreeding [Duthie and Reid *in press*; @Penn2010]. Similarly, females might choose to avoid inbreeding, but be ecologically constrained in their choice by the availability of males if few non-relatives are available [e.g., @Kokko2006a; @Kokko2013], or if females are unable to resist male coercion given sexual conflict [@Arnqvist2005, pp. 44-91]. Females might also be unable to discriminate between sperm of relatives versus non-relatives, or be unable to successfully bias fertilisation toward non-relative's sperm. Our results suggest that such constraints on one strategy of inbreeding avoidance (pre-copulatory versus post-copulatory) will strengthen selection on a complementary strategy, perhaps making inbreeding avoidance more detectable in empirical systems.

>>>>

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

We first test our hypothesis that a focal female's ability to express post-copulatory choice with respect to kinship causes selection for polyandry by comparing simulations in which post-copulatory inbreeding strategy is fixed to random mating with respect to kinship to simulations in which tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy can all evolve. For low costs of tendency for polyandry, $c_{P} < 0.005$, selection caused expected tendency for polyandry allele values ($P_{a}$) to be positive only when post-copulatory inbreeding avoidance evolved (Figure 1A-D), and for all $c_{P}$, $P_{a}$ values were higher when post-copulatory inbreeding avoidance evolved than when it could not evolve (Figure 1). When post-copulatory inbreeding avoidance could not evolve, $P_{a}$ values were negative even in the absence of a direct cost ($c_{P}=0$; Figure 1A) because females that expressed polyandry were more likely to increase their inbreeding due to sampling males without replacement [@Duthie]. Strong post-copulatory inbreeding avoidance consistently evolved in all simulations, even when $P_{a}$ values were not expected to be positive (Figure 1F,H), because of the threshold nature of the polyandry phenotype; Figure 2 shows how phenotypic polyandry was common even when mean polyandry allele values within a populatoin were expected to be negative. In all simulations, pre-copulatory inbreeding avoidance also evolved ($M_{a}<0$), meaning that both pre- and post-copulatory inbreeding avoidance was predicted, and evolution of adaptive polyandry occurred given sufficiently low costs of tendency for polyandry.

@@

Our second hypothesis that selection for a costly inbreeding avoidance strategy will be negligible given a less costly alternative was generally supported by our simulations (Figure 3). For simulations in which either pre- or post-copulatory inbreeding strategy allele values were fixed for random choice with respect to kinship ($M_{a}=0$ or $F_{a}=0$, respectively), the alternative costly inbreeding strategy evolved to inbreeding avoidance ($F_{a}<0$ or $M_{a}<0$). However, when the alternative inbreeding strategy could evolve and was not costly, the focal costly inbreeding strategy evolved toward random choice given increasingly negative allele values of the alternative strategy over generations (Figure 3B,D). For example, given fixed $F_{a}=0$, grand mean $M_{a}$ values were $-0.868$ in the last $10000$ generations of evolution (Figure 3A), meaning that an outbred focal female was ca $5.34$ times less likely to choose a full-sibling and ca $2.10$ times less likely to choose a first cousin than a non-relative given a strong cost of pre-copulatory inbreeding avoidance, $c_{M}=0.02$. In contrast, grand mean $M_{a}$ allele values were $-0.288$ when post-copulatory inbreeding avoidance evolved. More strongly, given fixed $M_{a}=0$, grand mean $F_{a}$ values were $-1.092$ in the last $10000$ generations of evolution (Figure 3C), meaning that an outbred focal female was ca $6.46$ times less likely to choose a full-sibling and ca $2.37$ times less likely to choose a first cousin than a non-relative given a strong cost of pre-copulatory inbreeding avoidance, $c_{F}=0.02$. However, grand-mean $F_{a}$ values were $-0.070$ when pre-copulatory inbreeding strategy evolved, and polyandry allele values were negative. Overall, selection for a costly inbreeding strategy that was adaptive in isolation was greatly weakened given evolution of a less costly strategy.

<!--- Figure 1 shows changes in mean allele values of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy across all eight possible combinations of costs. In Figure 1A, when no phenotypes are costly, inbreeding mating strategy and inbreeding fertilisation strategy evolve to inbreeding avoidance ($M_{a}<0$ and $F_{a}<0$), and females evolve to be polyandrous ($P_{a}>0$). Costs of phenotypes had a direct effect on evolution, causing inbreeding mating strategy and inbreeding fertilisation strategy phenotypes to evolve to values closer to zero (because realisation of inbreeding strategy costs increases with phenotype absolute values) and tendency for polyandry phenotypes to evolve to values that were more negative (because realisation of polyandry costs increases with positive phenotype values). --->

<!--- We can use Figure 1 to test the hypothesis that inbreeding mating strategy will evolve to inbreeding avoidance ($M_{a}<0$), but inbreeding fertilisation strategy will be negligible ($F_{a} \approx 0$) when mating costs are low ($c_{M}=0$) relative to the costs of tendency for polyandry and inbreeding fertilisation strategy ($c_{P}=0.02$ or $c_{F}=0.02$) by comparing different cost combinations to the null model in which all costs equal zero (Figure 1A). Figure 1A,C,D,G show that inbreeding avoidance in mating ($M_{a}<0$) evolved consistently whenever $c_{M}=0$. When inbreeding fertilisation strategy was costly and inbreeding avoidance strategy was not costly (Figure 1D,G), inbreeding fertilisation strategy initially evolved toward inbreeding avoidance ($F_{a} < 0$), but long-term evolution of inbreeding fertilisation strategy stabilised at random mating ($F_{a} \approx 0$). However, when only tendency for polyandry was costly (Figure 1C), both inbreeding mating strategy and inbreeding fertilisation strategy evolved to inbreeding avoidance over generations ($M_{a} < 0$ and $F_{a} < 0$). --->

<!--- We can also use Figure 1 to test the hypothesis that inbreeding fertilisation strategy will evolve to inbreeding avoidance ($F_{a}<0$), but inbreeding mating strategy will be negligible ($M_{a} \approx 0$) when tendency for polyandry and inbreeding fertilisation strategy costs are low ($c_{P}=0$ and $c_{F}=0$) relative to the cost of inbreeding mating strategy ($c_{M}=0.02$). Figure 1B shows that inbreeding avoidance through fertilisation bias evolved ($F_{a}<0$) under these conditions. While inbreeding mating strategy initially evolved toward inbreeding avoidance ($M_{a} < 0$), long-term evolution of inbreeding mating strategy stabilised at random mating ($M_{a} \approx 0$). --->

@@

Our third hypothesis that invasion of pre- or post-copulatory inbreeding avoidance would be precluded by the existence of an already adaptive complementary strategy was supported for invading post-copulatory, but not pre-copulatory, inbreeding avoidance. When pre-copulatory inbreeding strategy was initialised for random mating, but polyandry and post-copulatory inbreeding strategy allele values were fixed for adaptive inbreeding avoidance (Figure 4A,B), negative pre-copulatory inbreeding allele values became increasingly more freqent ($M_{a}<0$), causing inbreeding avoidance in mate choice. However, $M_{a}$ values were not as negative after $40000$ generations when post-copulatory inbreeding avoidance and polyandry were fixed as they were when no post-copulatory inbreeding avoidance occurred (e.g., compare Figure 4A with Figure 1A, and Figure 4B with Figure 2A). When post-copulatory inbreeding strategy was initialised for random fertilisation, existence of an already adaptive strategy of pre-copulatory inbreeding avoidance precluded evolution of polyandry and post-copulatory inbreeding avoidance (Figure 4C,D). Mean allele values affecting polyandry consistently decreased over generations regardless of the cost of post-copulatory inbreeding strategy, $c_{F}$. Consequently, when $c_{F}=0$, post-copulatory inbreeding strategy allele values had no effect because females were consistently monandrous, resulting in high drift of $F_{a}$ values among replicates (Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs.

<!--- Figure 4 shows change in inbreeding mating strategy allele values over generations when tendency for polyandry and inbreeding fertilisation strategy alleles were fixed for adaptive values of inbreeding avoidance $P_{a}=1$ and $F_{a}=-5$. Despite that females in these simulations could already avoid inbreeding after mating occurred, inbreeding mating strategy evolved towards inbreeding avoidance given costs of $c_{M}=0$ (Figure 4A) and $c_{M}=0.02$ (Figure 4B). We therefore did not find evidence that the existence of adaptive inbreeding avoidance through fertilisation bias in any way precluded the evolution of adaptive inbreeding avoidance through mate choice. --->

<!--- Figure 5 shows change in tendency for polyandry and inbreeding fertilisation strategy allele values over generations when inbreeding mating strategy alleles were fixed for adaptive inbreeding avoidance $M_{a}=-5$. Existence of adaptive inbreeding avoidance through mate choice precluded evolution of polyandry for all combinations of costs (Figure 5), meaning that evolution of adpative polyandry to avoid inbreeding through inbreeding fertilisation strategy was never predicted in populations where females were already avoiding inbreeding through mate choice. Further, while inbreeding fertilisation strategy evolved toward inbreeding avoidance ($F_{a}<0$) in the absence of costs (Figure 5A,B), allele values of inbreeding fertilisation strategy were not as negative as when adaptive inbreeding avoidance through mate choice was not fixed at the outset (Figures 1A,B,C,D). We therefore found strong evidence that the existence of adaptive inbreeding avoidance through pre-copulatory mate choice precluded evolution of adaptive inbreeding avoidance through post-copulatory inbreeding avoidance through fertilisation bias. --->

@@

<!--- Inbred offspring are typically less viable than outbred offspring, and when such inbreeding depression is strong, it is predicted to cause selection for inbreeding avoidance [Duthie and Reid *in press*; @Parker1979; @Parker2006; @Pusey1996; @Kokko2006]. In animals, inbreeding avoidance can be realised through both pre-copulatory and post-copulatory mechanisms [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, the conditions under which inbreeding avoidance is predicted to evolve via pre-copulatory, post-copulatory, or both pre- and post-copulatory mechanisms has not been modelled. There is therefore no framework for predicting what mechanisms of inbreeding avoidance are most likely to evolve in empirical populations where risk of fitness loss caused by inbreeding is high. Importantly, evolution of post-copulatory mechanisms of inbreeding avoidance requires females to mate multiply, meaning that post-copulatory inbreeding avoidance cannot occur unless polyandry has evolved in a population. Accordingly, --->

We modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency to engage in polyandry, inbreeding mating strategy (pre-copulatory), and inbreeding fertilisation strategy (post-copulatory). Using *in silico* experiments, we tested three specific hypothesis to predict evolution of polyandry, and pre- and post-copulatory inbreeding strategy. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of post-copulatory inbreeding avoidance, (2) selection for costly pre- or post-copulatory inbreeding avoidance was negligible given evolution of a less costly strategy of inbreeding avoidance, and (3) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of post-copulatory inbreeding avoidance, but not vice versa. More generally, we provide a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms.

<!--- Evolution of polyandry to avoid inbreeding --->
<!--- ---------------------------------------------- --->

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry and pre-copulatory inbreeding avoidance is predicted to be highly restrictive, requiring strong constraints on the availability of initial compared to additional mates, or the conditional expression of polyandry such that females only engaged in polyandry when their initial choice of mate was of low quality. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express post-copulatory choice. In contrast, we found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g. @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. For example, @Pizzari2004 found that the number of sperm arriving to eggs is lower for full siblings of female red jungefowl (*Gallus gallus*) than it is for non-relatives, even after controlling for effects of order of oviposition and social familiarity. Similarly, Firman and Simmons [-@Firman2008; -@Firman2015] have shown evidence of post-copulatory inbreeding avoidance in house mice, finding fertilisation to be biased toward unrelated males independent of the order in which mating occurs [@Firman2008], perhaps due to post-copulatory mechanisms expressed through the secretion of gametic proteins originating in female's eggs [@Firman2015]. Interestingly, female house mice have also been shown to express pre-copulatory inbreeding avoidance  [@Potts1991; @Roberts2003]. Our results suggest that post-copulatory inbreeding avoidance is unlikely to evolve when females can already avoid inbreeding through pre-copulatory mechanisms (Figure 4), meaning that evolution of post-copulatory inbreeding avoidance could reasonably be hypothesised to have preceded evolution of pre-copulatory inbreeding avoidance in mice.

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found that female *T. commodus* are capable of controlling the transfer of sperm to their spermatheca, most likely through physical control of the spermathecal duct, and can reduce sperm storage despite being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

<!---Costs of polyandry and inbreeding avoidance --->
<!--- ---------------------------------------------- --->

<!---- Our results show that direct negative selection on phenotypes that increase female promiscuity or choosiness with respect to kinship decrease the strength of selection for tendency for polyandry and inbreeding avoidance (Figure 1). Importantly, the cost of tendency for polyandry only decreased female fitness given tendency for polyandry phenotype values greater than zero ($P_{p}>0$), but the threshold nature of the tendency for polyandry phenotype only resulted in selection against positive tendency for polyandry allele values ($P_{a}$) when the sum over a focal female's $20$ $P_{a}$ alleles was greater than zero because summations less than zero all resulted in equivalent monandrous phenotypes ($P_{p}=0$). Hence, even when tendency for polyandry was costly, selection for more negative $P_{a}$ values weakened as the frequency of negative $P_{a}$ values increased because each individual $P_{a}$ then became less influential in causing females to be monandrous [@Duthie]. This general property of threshold traits causes maladaptive phenotypes to frequently reappear within populations due to the mutation and recombination of continuous genetic values underlying discrete phenotypes [@Roff1996; @Roff1998]. Assuming that there is continuous genetic variation underlying threshold polyandry, polyandrous phenotypes are therefore sometimes predicted to persist within populations even when costly [@Duthie], potentially creating the possibility for selection on inbreeding fertilisation strategy when polyandry is expressed. This is most evident in Figure 1C, where mean allele values for polyandry are negative (red line), but strong post-copulatory inbreeding avoidance is still predicted (grey line). Consequently, costly polyandry does not necessarily preclude evolution of adaptive post-copulatory inbreeding avoidance in populations. --->

In our model, a cost free inbreeding strategy precluded a more costly complementary inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre- and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of tendency for polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of inbreeding mating strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of inbreeding fertilisation strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations. <!--- add Tuni citation for depletion of energy? --->

@@

<!--- Constraints on polyandry and inbreeding avoidance --->
<!--- ---------------------------------------------- --->

@@

<!--- Constraints on pre-copulatory and post-copulatory inbreeding avoidance might vary widely in empirical systems. Female mate choice might be evolutionarily constrained if females are unable to discriminate among kin and therefore incapable of choosing to avoid inbreeding [Duthie and Reid *in press*; @Penn2010]. Similarly, females might choose to avoid inbreeding, but be ecologically constrained in their choice by the availability of males if few non-relatives are available [e.g., @Kokko2006a; @Kokko2013], or if females are unable to resist male coercion given sexual conflict [@Arnqvist2005, pp. 44-91]. Females might also be unable to discriminate between sperm of relatives versus non-relatives, or be unable to successfully bias fertilisation toward non-relative's sperm. Our results suggest that such constraints on one strategy of inbreeding avoidance (pre-copulatory versus post-copulatory) will strengthen selection on a complementary strategy, perhaps making inbreeding avoidance more detectable in empirical systems.  --->
